---
title: " Phylogenetic congruence between Neotropical primates and plants is driven by frugivory"
author: "L Fuzessy, FAO Silveira, L Culot, P Jordano & M Verdú"
date: "11/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load packages

```{r load.packs, include=FALSE}
library(phytools)
library(igraph)
library(ape)
library(picante)
library(dplyr)
library(paco)
library(ggplot2)

```

# Input interaction dataset 

```{r data.input, include=FALSE}

# Raw interaction dataset for Atlantic network.
int_data<- 
  read.table("./interaction_data.txt", 
             na="NA", sep="\t", dec=".", header= TRUE)


```


# Edge lists to interaction matrices

## Genus level
```{r edge.to.interact.gen, include=FALSE}

# Interaction datasets.


int_data %>%                      # Primates: 39 species
  dplyr::select(primate) %>%
  group_by(primate) %>%
  dplyr::mutate(count = n()) %>%
  unique()

int_data %>%                      # Plants: 312 genus
  dplyr::select(genus) %>%
  group_by(genus) %>%
  dplyr::mutate(count = n()) %>%
  unique()


# List of distinct pairwise interactions (edge list).
el_int_data.gen<- int_data %>%
  dplyr::select(primate,genus) %>%
  dplyr::filter(!is.na(primate)) %>%
  dplyr::group_by(primate,genus) %>%
  dplyr::summarise(n_distinct(primate))

```

## Species level
```{r edge.to.interact.spp, include=FALSE}

# Interaction datasets.


int_data %>%                      # Primates: 39 species
  dplyr::select(primate) %>%
  group_by(primate) %>%
  dplyr::mutate(count = n()) %>%
  unique()

int_data %>%                      # Plants: 964 plant
  dplyr::select(plant) %>%
  group_by(plant) %>%
  dplyr::mutate(count = n()) %>%
  unique()

# List of distinct pairwise interactions (edge list).
el_int_data.sp<- int_data %>%
  dplyr::select(primate,plant) %>%
  dplyr::filter(!is.na(primate)) %>%
  dplyr::group_by(primate,plant) %>%
  dplyr::summarise(n_distinct(primate))



```


# Create djacency matrices

## Plants at genus level

```{r adj.mat.gen, include=FALSE}

# Graph from edge list
int_data_g<- igraph::graph.data.frame(el_int_data.gen[,1:2], directed=FALSE)
V(int_data_g)$type <- V(int_data_g)$name %in% el_int_data.gen[,1]
is.bipartite(int_data_g)

int_data.mat.gen <- igraph::as_adjacency_matrix(int_data_g, type="both", names=T, sparse=F)

#Subset the adjacency matrix: primates as columns, plants as rows
int_data.mat.gen<- data.frame(int_data.mat.gen[1:39,40:351])
#write.csv(int_data.mat.gen, file= "interactions_gen.csv")
#Average number of interactions per species
(I.gen=sum(int_data.mat.gen!=0)) #Total number of interactions = 1060
(I.animals.gen=I.gen/nrow(int_data.mat.gen)) #26.18
(I.plants.gen=I.gen/ncol(int_data.mat.gen)) #3.40
```

## Plant at species level

```{r adj.mat.spp, include=FALSE}
# Graph from edge list
int_data_sp<- igraph::graph.data.frame(el_int_data.sp[,1:2], directed=FALSE)
V(int_data_sp)$type <- V(int_data_sp)$name %in% el_int_data.sp[,1]
is.bipartite(int_data_sp)

int_data.mat.sp <- igraph::as_adjacency_matrix(int_data_sp, type="both", names=T, sparse=F)

int_data.mat.sp <- igraph::as_adjacency_matrix(int_data_sp, type="both", names=T, sparse=F)

# Subset the adjacency matrix: primates as columns, plants as rows
int_data.mat.sp<- data.frame(int_data.mat.sp[1:39,40:1003])
#write.csv(int_data.mat.sp, file= "interactions_sp.csv")
#Average number of interactions per species
(I.sp=sum(int_data.mat.sp!=0)) #Total number of interactions = 1632
(I.animals.sp=I.sp/nrow(int_data.mat.sp)) #41.84
(I.plants.sp=I.sp/ncol(int_data.mat.sp)) #1.69

```


# Load trees

```{r load.trees, echo=TRUE}

prim_tree <- ape::read.nexus("./filtered_prim_tree.nex")

plant_tree.sp <- ape::read.nexus("./filtered_plant_tree_sp.nex")

plant_tree.gen <- ape::read.nexus("./filtered_plant_tree_gen.nex")

```



# Filtering interactions

## Genus level
```{r filter.genus}
interactions<-int_data.mat.gen
filtered_interactions.gen<-interactions[which(rownames(interactions) %in%
prim_tree$tip.label), which(colnames(interactions) %in%
plant_tree.gen$tip.label)]

filtered_prim_tree<-drop.tip(prim_tree, setdiff(prim_tree$tip.label,
rownames(filtered_interactions.gen)))
plot.phylo(filtered_prim_tree,  cex = 0.6, edge.width = 0.5)
##write.nexus(filtered_prim_tree, file="filtered_prim_tree.nex")

filtered_plant_tree.gen<-drop.tip(plant_tree.gen, setdiff(plant_tree.gen$tip.label,
colnames(filtered_interactions.gen)))
plot.phylo(filtered_plant_tree.gen, cex = 0.6, edge.width = 0.5)
##write.nexus(filtered_plant_tree, file="filtered_plant_tree.nex")
```

## Species level
```{r filter.species}

interactions.sp<-int_data.mat.sp
filtered_interactions.sp<-interactions.sp[which(rownames(interactions.sp) %in%
prim_tree$tip.label), which(colnames(interactions.sp) %in%
plant_tree.sp$tip.label)]

filtered_prim_tree<-drop.tip(prim_tree, setdiff(prim_tree$tip.label,
rownames(filtered_interactions.sp)))
plot.phylo(filtered_prim_tree,  cex = 0.6, edge.width = 0.5, type= "fan")
#write.nexus(filtered_prim_tree.sp, file="filtered_prim_tree_sp.nex")

filtered_plant_tree.sp<-drop.tip(plant_tree.sp, setdiff(plant_tree.sp$tip.label,
colnames(filtered_interactions.sp)))
plot.phylo(filtered_plant_tree.sp, cex = 0.6, edge.width = 0.5, type= "fan")
#write.nexus(filtered_plant_tree.sp, file="filtered_plant_tree_sp.nex")
```

# Perform PACo analysis

To perform PACo, the symmetric argument will be set to TRUE.
This means that both phylogenies (frugivores and plants) will be standardised prior to super-imposition resulting in the best-fit of the superimposition being independent of both phylogenies.

Also, as seed dispersal is a more diffuse ecological system (such as pollination), we will implement a more constrained null model (method=‘quasiswap’), where the number of interactions is conserved for each species and accordingly in the network as a whole.

# Genus level
```{r paco.genus}
D.gen<-prepare_paco_data(
      cophenetic(filtered_prim_tree),
      cophenetic(filtered_plant_tree.gen), 
      as.matrix(filtered_interactions.gen)
      )

# cophylogenetic analysis

D.gen <- add_pcoord(D.gen)
D.gen <- PACo(D.gen, nperm=30, method='quasiswap', symmetric=TRUE) #set nperm to 1000
D.gen$gof #p = 0, ss = 0.9842524, n = 1000

# investigate the contribution of individual links
res.gen <- residuals_paco(D.gen$proc)

```

## Species level
```{r paco.sp}

D.sp<-prepare_paco_data(
      cophenetic(filtered_prim_tree),
      cophenetic(filtered_plant_tree.sp), 
      as.matrix(filtered_interactions.sp)
      )

# cophylogenetic analysis
D.sp <- add_pcoord(D.sp)

# Pull vectors out of object D for PACo to read
X<-D.sp$H_PCo
Y<-D.sp$P_PCo

# cophylogenetic analysis
D.sp <- PACo(D.sp, nperm=20, method='quasiswap', symmetric=TRUE) #set nperm to 1000
#Check for deep errors produced by complex numbers. Use file 'coordpcoa2.r' - correction by Miguel

D.sp$gof #p = 0, ss = 0.97869, n = 1000

res.sp=read.table("./ele13918-sup-0002-TableS9.txt", 
             na="NA", sep="\t", dec=".", header= TRUE) #supp material

```


# Visualize partitioned residual contribuiton

## Genus level
```{r resid.cont.gen}

# to analyse the links splitting the interactions based on the plot
#first, selecting the most frugivorous)
cophy_int_fru <- c(grep('Ateles_belzebuth', names(res.gen)), 
               grep('Ateles_geoffroyi', names(res.gen)),
               grep('Ateles_paniscus', names(res.gen)),
               grep('Brachyteles_arachnoides', names(res.gen)),
               grep('Brachyteles_hypoxanthus', names(res.gen)),
               grep('Lagothrix_lagotricha', names(res.gen)),
               grep('Lagothrix_cana', names(res.gen)),
               grep('Lagothrix_lugens', names(res.gen))
               )
#first, selecting the most folivorous
cophy_int_fol <- c(grep('Alouatta_belzebul', names(res.gen)), 
               grep('Alouatta_caraya', names(res.gen)), 
               grep('Alouatta_guariba', names(res.gen)), 
               grep('Alouatta_palliata', names(res.gen)), 
               grep('Alouatta_pigra', names(res.gen)), 
               grep('Alouatta_seniculus', names(res.gen))
               )

#second, selecting the 'insectivorous'
cophy_int_ins <- c( 
               grep('Leontopithecus_caissara', names(res.gen)), 
               grep('Leontopithecus_chrysomelas', names(res.gen)),
               grep('Leontopithecus_chrysopygus', names(res.gen)),
               grep('Leontopithecus_rosalia', names(res.gen)),
               grep('Saguinus_fuscicollis', names(res.gen)),
               grep('Saguinus_geoffroyi', names(res.gen)),
               grep('Saguinus_midas', names(res.gen)),
               grep('Saguinus_mystax', names(res.gen)),
              grep('Callithrix_aurita', names(res.gen)),
	            grep('Callithrix_flaviceps', names(res.gen)),
	          	grep('Callithrix_geoffroyi', names(res.gen)),
	          	grep('Callithrix_jacchus', names(res.gen)),
	          	grep('Callithrix_penicillata',names(res.gen))
                )
#second, selecting the Omnivorous               
cophy_int_omn <- c(grep('Sapajus_apella', names(res.gen)), 
               grep('Sapajus_flavius', names(res.gen)), 
               grep('Sapajus_libidinosus', names(res.gen)), 
               grep('Sapajus_nigritus', names(res.gen)), 
               grep('Sapajus_robustus', names(res.gen)), 
               grep('Sapajus_xanthosternos', names(res.gen)), 
               grep('Cebus_capucinus', names(res.gen)) 
               )

#Finally, selecting Pitheciidae (the seed predators)
cophy_int_pre <- c(grep('Cacajao_melanocephalus', names(res.gen)), 
               grep('Callicebus_coimbrai', names(res.gen)), 
               grep('Callicebus_melanochir', names(res.gen)), 
               grep('Callicebus_nigrifrons', names(res.gen)), 
               grep('Chiropotes_chiropotes', names(res.gen))
                )


# if need to remove the any interaction 
####cophy_int <- cophy_int[-grep('Anthidium', names(res[cophy_int]))]

# Visualise residuals of each primate family and the total
cophy_fru <- res.gen[cophy_int_fru]
cophy_fol <- res.gen[cophy_int_fol]
cophy_ins <- res.gen[cophy_int_ins]
cophy_omn <- res.gen[cophy_int_omn]
cophy_pre <- res.gen[cophy_int_pre]

# Save as data frame (ggplot do not run if it is not a data frame)
cophy_fru_dat <- as.data.frame(cophy_fru) 
cophy_fol_dat <- as.data.frame(cophy_fol) 
cophy_ins_dat <- as.data.frame(cophy_ins) 
cophy_omn_dat <- as.data.frame(cophy_omn)
cophy_pre_dat <- as.data.frame(cophy_pre) 
  


# visualise the difference with a box and whisker plot

dat<- rbind(
        data.frame(cophy=cophy_fru, level='high', diet='fru'), 
        data.frame(cophy=cophy_fol, level='rest',diet='fol'),
        data.frame(cophy=cophy_omn,level='rest', diet='omn'),
        data.frame(cophy=cophy_ins,level='rest', diet='ins'),
        data.frame(cophy=cophy_pre, level='rest',diet='pre'))

 ggplot(dat, aes(x=diet, y=cophy))+
  geom_boxplot(alpha=0.85, fill=c('white'), outlier.color ='white')+
  geom_jitter(data=dat, aes(x=diet, y=cophy, col=diet), alpha=0.4)+
  scale_fill_brewer(palette='Paired')+
  scale_x_discrete(labels=c('Frugivore','Folivore', 'Omnivore', 'Insectivore', 'Seed Predator'))+
  ylab('Contribution to cophylogenetic signal')+
  theme_bw()+
  theme(
    axis.title.x=element_blank(),
    panel.grid=element_blank(),
    legend.position='none',
    axis.text.x=element_text(size=12),
    axis.title.y=element_text(size=14)
  )
 

 
 
```

## Species level
```{r resid.cont.sp}

# to analyse the links splitting the interactions based on the plot
#first, selecting the most frugivorous)
cophy_int_fru <- res.sp %>%
  filter(primate_spp %in% c('Ateles_belzebuth', 
               'Ateles_geoffroyi',
               'Ateles_paniscus',
               'Brachyteles_arachnoides',
               'Brachyteles_hypoxanthus',
               'Lagothrix_lagotricha',
               'Lagothrix_cana',
               'Lagothrix_lugens'
               ))

#first, selecting the most folivorous
cophy_int_fol <- res.sp %>%
  filter(primate_spp %in% c('Alouatta_belzebul',
                            'Alouatta_caraya',
                            'Alouatta_guariba',
                            'Alouatta_palliata',
                            'Alouatta_pigra', 
                            'Alouatta_seniculus'))

#second, selecting the 'insectivorous'
cophy_int_ins <- res.sp %>%
  filter(primate_spp %in% c('Leontopithecus_caissara',  
               'Leontopithecus_chrysomelas', 
               'Leontopithecus_chrysopygus', 
               'Leontopithecus_rosalia', 
               'Saguinus_fuscicollis', 
               'Saguinus_geoffroyi', 
               'Saguinus_midas', 
               'Saguinus_mystax', 
              'Callithrix_aurita', 
	            'Callithrix_flaviceps', 
	          	'Callithrix_geoffroyi', 
	          	'Callithrix_jacchus', 
	          	'Callithrix_penicillata')
                )
#second, selecting the Omnivorous               
cophy_int_omn <- res.sp %>%
  filter(primate_spp %in% c('Sapajus_apella',  
               'Sapajus_flavius',  
               'Sapajus_libidinosus',  
               'Sapajus_nigritus',  
               'Sapajus_robustus',  
               'Sapajus_xanthosternos',  
               'Cebus_capucinus') 
               )

#Finally, selecting Pitheciidae (the seed predators)
cophy_int_pre <- res.sp %>%
  filter(primate_spp %in% c('Cacajao_melanocephalus',  
               'Callicebus_coimbrai',  
               'Callicebus_melanochir',  
               'Callicebus_nigrifrons',  
               'Chiropotes_chiropotes')
                )


# Save as data frame (ggplot do not run if it is not a data frame)
cophy_fru_dat <- as.data.frame(cophy_int_fru) 
cophy_fol_dat <- as.data.frame(cophy_int_fru) 
cophy_ins_dat <- as.data.frame(cophy_int_fru) 
cophy_omn_dat <- as.data.frame(cophy_int_fru)
cophy_pre_dat <- as.data.frame(cophy_int_fru) 
  


# visualise the difference with a box and whisker plot

dat<- rbind(
        data.frame(cophy=cophy_fru, level='high', diet='fru'), 
        data.frame(cophy=cophy_fol, level='rest',diet='fol'),
        data.frame(cophy=cophy_omn,level='rest', diet='omn'),
        data.frame(cophy=cophy_ins,level='rest', diet='ins'),
        data.frame(cophy=cophy_pre, level='rest',diet='pre'))

 ggplot(dat, aes(x=diet, y=cophy))+
  geom_boxplot(alpha=0.85, fill=c('white'), outlier.color ='white')+
  geom_jitter(data=dat, aes(x=diet, y=cophy, col=diet), alpha=0.4)+
  scale_fill_brewer(palette='Paired')+
  scale_x_discrete(labels=c('Frugivore','Folivore', 'Omnivore', 'Insectivore', 'Seed Predator'))+
  ylab('Contribution to cophylogenetic signal')+
  theme_bw()+
  theme(
    axis.title.x=element_blank(),
    panel.grid=element_blank(),
    legend.position='none',
    axis.text.x=element_text(size=12),
    axis.title.y=element_text(size=14)
  )
 

 
 
```


# Explore sampling effects
Better studied monkeys are those who contribute more to the cophylogenetic signal?

## Species level
```{r sample.effect}

sample<-read.table("./ref_interact2.txt", header= TRUE)

ref_sites<-read.table("./ref_sites.txt", 
                   na="NA", sep="\t", dec=".", header= TRUE)

sample_refs=left_join (sample,ref_sites, by='ref')
sample_refs_sp=sample_refs [!duplicated(sample_refs[c(1,2,3)]),]
sample<-read.table("./effort.txt", 
                   na="NA", sep="\t", header= TRUE)


#First, calculate the average Procrustes res.spidual across plant genus
res.spiduals=as.data.frame(res.sp)
pl_res.spid <- as.data.frame(res.spiduals[,c(2,3)])

library(plyr)
pl_res.sp_mean = ddply(pl_res.spid, .(plant_spp),transform, mean_res.sp = mean(res))
#Then, calculate the average Procrustes residual across primate species
pr_res.spid <- as.data.frame(res.spiduals[,c(1,3)])
pr_res.sp_mean = plyr::ddply(pr_res.spid,.(primate_spp),transform,mean_res.sp = mean(res))

names(pr_res.sp_mean)[names(pr_res.sp_mean) == "primate_spp"] <- "prim"
names(pl_res.sp_mean)[names(pr_res.sp_mean) == "plant_spp"] <- "plant"

#Join average Procrustes residual for each primate species with sample effort
sample_refs=left_join (pr_res.sp_mean,sample, by='prim')
sample_refs=sample_refs[,c(1,3,4)]
sample_refs=unique(sample_refs)



p=ggplot(sample_refs, aes(x=effort_h, y=mean_res.sp))+
  geom_smooth(method=lm, 
              color="darkred", fill="darkred", se=TRUE, alpha=0.4) +
  ylab('Average Procrustes res.spidual')+
  xlab('Sampling effort (h)')+
  theme_bw()+
  theme(
    panel.grid=element_blank(),
    legend.position='none',
    axis.title.x=element_text(size=14),
    axis.title.y=element_text(size=14)
  )
p

```



# Explore 'Local' patterns

## Subset data according to region 

```{r subset.region}
forest_interact=read.table("./interaction_forest.txt", 
             na="NA", sep="\t", dec=".", header= TRUE)


#subset by region
at_fo<- forest_interact %>%             
        dplyr::filter(forest=="at_fo")

am_fo<- forest_interact %>%             
        dplyr::filter(forest=="am_fo")

me_fo<- forest_interact %>%             
        dplyr::filter(forest=="me_fo")

```

## Edge lists to interaction matrices

```{r edge.to.int.local.spp, include=FALSE}

###Interaction datasets.

#ATLANTIC FOREST
at_fo %>%                      # primate: 22 species
  dplyr::select(primate) %>%
  group_by(primate) %>%
  dplyr::mutate(count = n()) %>%
  unique()

at_fo %>%                      # Plants: 382 plant
  dplyr::select(plant) %>%
  group_by(plant) %>%
  dplyr::mutate(count = n()) %>%
  unique()

#AMAZONIA
am_fo %>%                      # primate: 12 species
  dplyr::select(primate) %>%
  group_by(primate) %>%
  dplyr::mutate(count = n()) %>%
  unique()

am_fo %>%                      # Plants: 463 plant
  dplyr::select(plant) %>%
  group_by(plant) %>%
  dplyr::mutate(count = n()) %>%
  unique()


#MESOAMERICA FOREST
me_fo %>%                      # primate: 6 species
  dplyr::select(primate) %>%
  group_by(primate) %>%
  dplyr::mutate(count = n()) %>%
  unique()

me_fo %>%                      # Plants: 212 plant
  dplyr::select(plant) %>%
  group_by(plant) %>%
  dplyr::mutate(count = n()) %>%
  unique()

### List of distinct pairwise interactions (edge list).
el_at_fo.sp<- at_fo %>%
  dplyr::select(primate,plant) %>%
  dplyr::filter(!is.na(primate)) %>%
  dplyr::group_by(primate,plant) %>%
  dplyr::summarise(n_distinct(primate))

el_am_fo.sp<- am_fo %>%
  dplyr::select(primate,plant) %>%
  dplyr::filter(!is.na(primate)) %>%
  dplyr::group_by(primate,plant) %>%
  dplyr::summarise(n_distinct(primate))

el_me_fo.sp<- me_fo %>%
  dplyr::select(primate,plant) %>%
  dplyr::filter(!is.na(primate)) %>%
  dplyr::group_by(primate,plant) %>%
  dplyr::summarise(n_distinct(primate))


### Create graph objects
# Graph from edge list
at_fo_g.sp<- igraph::graph.data.frame(el_at_fo.sp[,1:2], directed=FALSE)
V(at_fo_g.sp)$type <- V(at_fo_g.sp)$name %in% el_at_fo.sp[,1]

am_fo_g.sp<- igraph::graph.data.frame(el_am_fo.sp[,1:2], directed=FALSE)
V(am_fo_g.sp)$type <- V(am_fo_g.sp)$name %in% el_am_fo.sp[,1]

me_fo_g.sp<- igraph::graph.data.frame(el_me_fo.sp[,1:2], directed=FALSE)
V(me_fo_g.sp)$type <- V(me_fo_g.sp)$name %in% el_me_fo.sp[,1]
```



## Create djacency matrices

```{r adj.mat.local.sp, include=FALSE}
at_fo.mat.sp <- igraph::as_adjacency_matrix(at_fo_g.sp, type="both", names=T, sparse=F)
# Subset the adjacency matrix: primates as columns, plants as rows
at_fo.mat.sp<- data.frame(at_fo.mat.sp[1:22,23:404])

am_fo.mat.sp <- igraph::as_adjacency_matrix(am_fo_g.sp, type="both", names=T, sparse=F)
# Subset the adjacency matrix: primates as columns, plants as rows
am_fo.mat.sp<- data.frame(am_fo.mat.sp[1:12,13:475])

me_fo.mat.sp <- igraph::as_adjacency_matrix(me_fo_g.sp, type="both", names=T, sparse=F)
# Subset the adjacency matrix: primates as columns, plants as rows
me_fo.mat.sp<- data.frame(me_fo.mat.sp[1:5,6:202])


```

### 1.16.c Filtering trees and interactions

```{r filter.tree.local.sp}

###ATLANTIC FOREST
#interactions<-int_data.mat.sp
filt_interact_at_fo.sp<-at_fo.mat.sp[which(rownames(at_fo.mat.sp) %in%
filtered_prim_tree$tip.label), which(colnames(at_fo.mat.sp) %in%
filtered_plant_tree.sp$tip.label)]

filt_prim_tree_at_fo.sp<-drop.tip(filtered_prim_tree, setdiff(filtered_prim_tree$tip.label,
rownames(filt_interact_at_fo.sp)))
plot.phylo(filt_prim_tree_at_fo.sp,  cex = 0.6, edge.width = 0.5)
##write.nexus(filtered_prim_tree.sp, file="filt_prim_tree_at_fo.nex")

filt_pl_tree_at_fo.sp<-drop.tip(filtered_plant_tree.sp, setdiff(filtered_plant_tree.sp$tip.label,
colnames(filt_interact_at_fo.sp)))
plot.phylo(filt_pl_tree_at_fo.sp, cex = 0.6, edge.width = 0.5)
##write.nexus(filtered_plant_tree.sp, file="filt_interact_at_fo.nex")

###AMAZON FOREST
#interactions<-int_data.mat.sp
filt_interact_am_fo.sp<-am_fo.mat.sp[which(rownames(am_fo.mat.sp) %in%
filtered_prim_tree$tip.label), which(colnames(am_fo.mat.sp) %in%
filtered_plant_tree.sp$tip.label)]

filt_prim_tree_am_fo.sp<-drop.tip(filtered_prim_tree, setdiff(filtered_prim_tree$tip.label,
rownames(filt_interact_am_fo.sp)))
#filtering for plant species with only 1 interaction 
filt_prim_tree_am_fo.sp<-drop.tip(filt_prim_tree_am_fo.sp, "Alouatta_palliata" )
filt_prim_tree_am_fo.sp<-drop.tip(filt_prim_tree_am_fo.sp, "Saimiri_ustus" )
filt_prim_tree_am_fo.sp<-drop.tip(filt_prim_tree_am_fo.sp, "Lagothrix_cana")
plot.phylo(filt_prim_tree_am_fo.sp,  cex = 0.6, edge.width = 0.5)
##write.nexus(filtered_prim_tree.sp, file="filt_prim_tree_am_fo.nex")

filt_pl_tree_am_fo.sp<-drop.tip(filtered_plant_tree.sp, setdiff(filtered_plant_tree.sp$tip.label,
colnames(filt_interact_am_fo.sp)))
plot.phylo(filt_pl_tree_am_fo.sp, cex = 0.6, edge.width = 0.5)
##write.nexus(filtered_plant_tree.sp, file="filt_interact_am_fo.nex")


###MESOAMERICAN FOREST
#interactions<-int_data.mat.sp
filt_interact_me_fo.sp<-me_fo.mat.sp[which(rownames(me_fo.mat.sp) %in%
filtered_prim_tree$tip.label), which(colnames(me_fo.mat.sp) %in%
filtered_plant_tree.sp$tip.label)]

filt_prim_tree_me_fo.sp<-drop.tip(filtered_prim_tree, setdiff(filtered_prim_tree$tip.label,
rownames(filt_interact_me_fo.sp)))
plot.phylo(filt_prim_tree_me_fo.sp,  cex = 0.6, edge.width = 0.5)
##write.nexus(filtered_prim_tree.sp, file="filt_prim_tree_me_fo.sp.nex")

filt_pl_tree_me_fo.sp<-drop.tip(filtered_plant_tree.sp, setdiff(filtered_plant_tree.sp$tip.label,
colnames(filt_interact_me_fo.sp)))
plot.phylo(filt_pl_tree_me_fo.sp, cex = 0.6, edge.width = 0.5)
##write.nexus(filtered_plant_tree.sp, file="filt_interact_me_fo.sp.nex")

```

# Perform PACo analysis - local

## ATLANTIC FOREST

### Species Level 

```{r paco.atlantic.forest.sp}
# start the paco procedure
filt_interact_at_fo.sp<-filt_interact_at_fo.sp[filt_prim_tree_at_fo.sp$tip.label,
filt_pl_tree_at_fo.sp$tip.label]

D_at.sp<-prepare_paco_data(
      cophenetic(filt_prim_tree_at_fo.sp),
      cophenetic(filt_pl_tree_at_fo.sp), 
      filt_interact_at_fo.sp
      )
D_at.sp <- add_pcoord(D_at.sp)

# cophylogenetic analysis
X<-D$H_PCo
Y<-D$P_PCo
procrustes(X,Y)
D_at.sp <- PACo(D_at.sp, nperm=1000, method='quasiswap', symmetric=TRUE) 

D_at.sp$gof #p=p-value >> 0, $ss = sum of residuals >> 0.982, $n=permutation number >>1000

# investigate the contribution of individual links
res_at.sp <- residuals_paco(D_at.sp$proc)
sort(res_at.sp,decreasing=TRUE)

```

Residual contribution visualization

```{r resid.contribut.at.fo.sp}


# to analyse the residuals based on all primates
cophy_at.sp=as.data.frame(res_at.sp)

# to analyse the links splitting the interactions based on the plot
#first, selecting the most frugivorous)
cophy_int_fru_at.sp <- c(grep('Brachyteles_arachnoides', names(res_at.sp)),
               grep('Brachyteles_hypoxanthus', names(res_at.sp))
               )
               
#first, selecting the most folivorous
cophy_int_fol_at.sp <- c(grep('Alouatta_belzebul', names(res_at.sp)), 
               grep('Alouatta_caraya', names(res_at.sp)), 
               grep('Alouatta_guariba', names(res_at.sp)), 
               grep('Alouatta_palliata', names(res_at.sp)), 
               grep('Alouatta_pigra', names(res_at.sp)), 
               grep('Alouatta_seniculus', names(res_at.sp))
               )

#second, selecting the 'insectivorous'
cophy_int_ins_at.sp <- c( 
               grep('Leontopithecus_caissara', names(res_at.sp)), 
               grep('Leontopithecus_chrysomelas', names(res_at.sp)),
               grep('Leontopithecus_chrysopygus', names(res_at.sp)),
               grep('Leontopithecus_rosalia', names(res_at.sp)),
               grep('Callithrix_aurita', names(res_at.sp)),
	            grep('Callithrix_flaviceps', names(res_at.sp)),
		grep('Callithrix_geoffroyi', names(res_at.sp)),
		grep('Callithrix_jacchus', names(res_at.sp)),
		grep('Callithrix_penicillata',names(res_at.sp))
                )
#second, selecting the Omnivorous               
cophy_int_omn_at.sp <- c(grep('Sapajus_apella', names(res_at.sp)), 
               grep('Sapajus_flavius', names(res_at.sp)), 
               grep('Sapajus_libidinosus', names(res_at.sp)), 
               grep('Sapajus_nigritus', names(res_at.sp)), 
               grep('Sapajus_robustus', names(res_at.sp)), 
               grep('Sapajus_xanthosternos', names(res_at.sp)) 
               )

#Finally, selecting Pitheciidae (the seed predators)
cophy_int_pre_at.sp <- c(grep('Cacajao_melanocephalus', names(res_at.sp)), 
               grep('Callicebus_coimbrai', names(res_at.sp)), 
               grep('Callicebus_melanochir', names(res_at.sp)), 
               grep('Callicebus_nigrifrons', names(res_at.sp)), 
               grep('Chiropotes_chiropotes', names(res_at.sp))
                )

# Visualise residuals of each primate family and the total
cophy_fru_at.sp <- res_at.sp[cophy_int_fru_at.sp]
cophy_fol_at.sp <- res_at.sp[cophy_int_fol_at.sp]
cophy_ins_at.sp <- res_at.sp[cophy_int_ins_at.sp]
cophy_omn_at.sp <- res_at.sp[cophy_int_omn_at.sp]
cophy_pre_at.sp <- res_at.sp[cophy_int_pre_at.sp]
   
dat_at.sp<- rbind(
        data.frame(cophy=cophy_fru_at.sp,  diet='1'), 
        data.frame(cophy=cophy_fol_at.sp, diet='2'),
        data.frame(cophy=cophy_omn_at.sp, diet='3'),
        data.frame(cophy=cophy_ins_at.sp, diet='4'),
        data.frame(cophy=cophy_pre_at.sp, diet='5'))

 ggplot(dat_at.sp, aes(x=diet, y=cophy))+
  geom_boxplot(alpha=0.85, fill=c('white'), outlier.color ='white')+
  geom_jitter(data=dat_at.sp, aes(x=diet, y=cophy, col=diet), alpha=0.4)+
  scale_fill_brewer(palette='Paired')+
  scale_x_discrete(labels=c('Frugivore','Folivore', 'Omnivore', 'Insectivore', 'Seed Predator'))+
  ylab('Contribution to cophylogenetic signal')+
  theme_bw()+
  theme(
    axis.title.x=element_blank(),
    panel.grid=element_blank(),
    legend.position='none',
    axis.text.x=element_text(size=12),
    axis.title.y=element_text(size=14)
  )
```

## AMAZON FOREST

### Species Level 

```{r paco.amazon.forest.sp}
# start the paco procedure
filt_interact_am_fo.sp<-filt_interact_am_fo.sp[filt_prim_tree_am_fo.sp$tip.label,
filt_pl_tree_am_fo.sp$tip.label]

D_am.sp<-prepare_paco_data(
      cophenetic(filt_prim_tree_am_fo.sp),
      cophenetic(filt_pl_tree_am_fo.sp), 
      filt_interact_am_fo.sp
      )
D_am.sp <- add_pcoord(D_am.sp)

# cophylogenetic analysis
D_am.sp <- PACo(D_am.sp, nperm=1000, method='quasiswap', symmetric=TRUE) 

D_am.sp$gof #p=p-value >> 0, $ss = sum of residuals >> 0.982, $n=permutation number >>1000

# investigate the contribution of individual links
res_am.sp <- residuals_paco(D_am.sp$proc)
sort(res_am.sp,decreasing=TRUE)




```

####Residual contribution visualization

```{r resid.contribut.am.fo.sp}


# to analyse the residuals based on all primates
cophy_am.sp=as.data.frame(res_am.sp)

# to analyse the links splitting the interactions based on the plot
#first, selecting the most frugivorous)
cophy_int_fru_am.sp <- c(grep('Ateles_belzebuth', names(res_am.sp)), 
               grep('Ateles_geoffroyi', names(res_am.sp)),
               grep('Ateles_paniscus', names(res_am.sp)),
               grep('Brachyteles_arachnoides', names(res_am.sp)),
               grep('Brachyteles_hypoxanthus', names(res_am.sp)),
               grep('Lagothrix_lagotricha', names(res_am.sp)),
               grep('Lagothrix_cana', names(res_am.sp)),
               grep('Lagothrix_lugens', names(res_am.sp))
               )
#first, selecting the most folivorous
cophy_int_fol_am.sp <- c(grep('Alouatta_belzebul', names(res_am.sp)), 
               grep('Alouatta_caraya', names(res_am.sp)), 
               grep('Alouatta_guariba', names(res_am.sp)), 
               grep('Alouatta_palliata', names(res_am.sp)), 
               grep('Alouatta_pigra', names(res_am.sp)), 
               grep('Alouatta_seniculus', names(res_am.sp))
               )

#second, selecting the 'insectivorous'
cophy_int_ins_am.sp <- c( 
               grep('Leontopithecus_caissara', names(res_am.sp)), 
               grep('Leontopithecus_chrysomelas', names(res_am.sp)),
               grep('Leontopithecus_chrysopygus', names(res_am.sp)),
               grep('Leontopithecus_rosalia', names(res_am.sp)),
               grep('Saguinus_fuscicollis', names(res_am.sp)),
               grep('Saguinus_geoffroyi', names(res_am.sp)),
               grep('Saguinus_midas', names(res_am.sp)),
               grep('Saguinus_mystax', names(res_am.sp))
                )
#second, selecting the Omnivorous               
cophy_int_omn_am.sp <- c(grep('Sapajus_apella', names(res_am.sp)), 
               grep('Sapajus_flavius', names(res_am.sp)), 
               grep('Sapajus_libidinosus', names(res_am.sp)), 
               grep('Sapajus_nigritus', names(res_am.sp)), 
               grep('Sapajus_robustus', names(res_am.sp)), 
               grep('Sapajus_xanthosternos', names(res_am.sp)), 
               grep('Cebus_capucinus', names(res_am.sp)) 
               )

#Finally, selecting Pitheciidae (the seed predators)
cophy_int_pre_am.sp <- c(grep('Cacajao_melanocephalus', names(res_am.sp)), 
               grep('Callicebus_coimbrai', names(res_am.sp)), 
               grep('Callicebus_melanochir', names(res_am.sp)), 
               grep('Callicebus_nigrifrons', names(res_am.sp)), 
               grep('Chiropotes_chiropotes', names(res_am.sp))
                )

# Visualise residuals of each primate family and the total
cophy_fru_am.sp <- res_am.sp[cophy_int_fru_am.sp]
cophy_fol_am.sp <- res_am.sp[cophy_int_fol_am.sp]
cophy_ins_am.sp <- res_am.sp[cophy_int_ins_am.sp]
cophy_omn_am.sp <- res_am.sp[cophy_int_omn_am.sp]
cophy_pre_am.sp <- res_am.sp[cophy_int_pre_am.sp]
   
dat_am.sp<- rbind(
        data.frame(cophy=cophy_fru_am.sp, diet='1'), 
        data.frame(cophy=cophy_fol_am.sp, diet='2'),
        data.frame(cophy=cophy_ins_am.sp, diet='3'),
        data.frame(cophy=cophy_pre_am.sp, diet='4'))

 ggplot(dat_am.sp, aes(x=diet, y=cophy))+
  geom_boxplot(alpha=0.85, fill=c('white'), outlier.color ='white')+
  geom_jitter(data=dat_am.sp, aes(x=diet, y=cophy, col=diet), alpha=0.4)+
    scale_colour_manual(values=c("#F8766D", "#93AA00", "#00B0F6","#C77CFF"))+
  scale_x_discrete(labels=c('Frugivore','Folivore', 'Insectivore', 'Seed Predator'))+
  ylab('Contribution to cophylogenetic signal')+
  theme_bw()+
  theme(
    axis.title.x=element_blank(),
    panel.grid=element_blank(),
    legend.position='none',
    axis.text.x=element_text(size=12),
    axis.title.y=element_text(size=14)
  )
```

###MESOAMERICAN FOREST

#### Species Level 

```{r paco.mesoam.forest.sp}
# start the paco procedure
filt_interact_me_fo.sp<-filt_interact_me_fo.sp[filt_prim_tree_me_fo.sp$tip.label,
filt_pl_tree_me_fo.sp$tip.label]

D_me.sp<-prepare_paco_data(
      cophenetic(filt_prim_tree_me_fo.sp),
      cophenetic(filt_pl_tree_me_fo.sp), 
      filt_interact_me_fo.sp
      )
D_me.sp <- add_pcoord(D_me.sp)

# cophylogenetic analysis
D_me.sp <- PACo(D_me.sp, nperm=1000, method='quasiswap', symmetric=TRUE) 

D_me.sp$gof #p=p-value >> 0, $ss = sum of residuals >> 0.982, $n=permutation number >>1000

# investigate the contribution of individual links
res_me.sp <- residuals_paco(D_me.sp$proc)
sort(res_me.sp,decreasing=TRUE)
#write.csv(res_me.sp, file="int_resid_me.sp.csv")



```

####Residual contribution visualization

```{r resid.contribut.me.fo.sp}


# to analyse the residuals based on all primates
cophy_me.sp=as.data.frame(res_me.sp)

# to analyse the links splitting the interactions based on the plot
#first, selecting the most frugivorous)
cophy_int_fru_me.sp <- c(grep('Ateles_belzebuth', names(res_me.sp)), 
               grep('Ateles_geoffroyi', names(res_me.sp)),
               grep('Ateles_paniscus', names(res_me.sp)),
               grep('Brachyteles_arachnoides', names(res_me.sp)),
               grep('Brachyteles_hypoxanthus', names(res_me.sp)),
               grep('Lagothrix_lagotricha', names(res_me.sp)),
               grep('Lagothrix_cana', names(res_me.sp)),
               grep('Lagothrix_lugens', names(res_me.sp))
               )
#first, selecting the most folivorous
cophy_int_fol_me.sp <- c(grep('Alouatta_belzebul', names(res_me.sp)), 
               grep('Alouatta_caraya', names(res_me.sp)), 
               grep('Alouatta_guariba', names(res_me.sp)), 
               grep('Alouatta_palliata', names(res_me.sp)), 
               grep('Alouatta_pigra', names(res_me.sp)), 
               grep('Alouatta_seniculus', names(res_me.sp))
               )

#second, selecting the 'insectivorous'
cophy_int_ins_me.sp <- c( 
               grep('Leontopithecus_caissara', names(res_me.sp)), 
               grep('Leontopithecus_chrysomelas', names(res_me.sp)),
               grep('Leontopithecus_chrysopygus', names(res_me.sp)),
               grep('Leontopithecus_rosalia', names(res_me.sp)),
               grep('Saguinus_fuscicollis', names(res_me.sp)),
               grep('Saguinus_geoffroyi', names(res_me.sp)),
               grep('Saguinus_midas', names(res_me.sp)),
               grep('Saguinus_mystax', names(res_me.sp))
                )
#second, selecting the Omnivorous               
cophy_int_omn_me.sp <- c(grep('Sapajus_apella', names(res_me.sp)), 
               grep('Sapajus_flavius', names(res_me.sp)), 
               grep('Sapajus_libidinosus', names(res_me.sp)), 
               grep('Sapajus_nigritus', names(res_me.sp)), 
               grep('Sapajus_robustus', names(res_me.sp)), 
               grep('Sapajus_xanthosternos', names(res_me.sp)), 
               grep('Cebus_capucinus', names(res_me.sp)) 
               )


# Visualise residuals of each primate family and the total
cophy_fru_me.sp <- res_me.sp[cophy_int_fru_me.sp]
cophy_fol_me.sp <- res_me.sp[cophy_int_fol_me.sp]
cophy_ins_me.sp <- res_me.sp[cophy_int_ins_me.sp]
cophy_omn_me.sp <- res_me.sp[cophy_int_omn_me.sp]
   
dat_me.sp<- rbind(
        data.frame(cophy=cophy_fru_me.sp, level='high', diet='1'), 
        data.frame(cophy=cophy_fol_me.sp, level='rest',diet='2'),
        data.frame(cophy=cophy_omn_me.sp,level='rest', diet='3'),
        data.frame(cophy=cophy_ins_me.sp,level='rest', diet='4'))

 ggplot(dat_me.sp, aes(x=diet, y=cophy))+
  geom_boxplot(alpha=0.85, fill=c('white'), outlier.color ='white')+
  geom_jitter(data=dat_me.sp, aes(x=diet, y=cophy, col=diet), alpha=0.4)+
   scale_colour_manual(values=c("#F8766D", "#A3A500", "#00BF7D", "#00B0F6"))+
  scale_x_discrete(labels=c('Frugivore','Folivore', 'Omnivore', 'Insectivore'))+
  ylab('Contribution to cophylogenetic signal')+
  theme_bw()+
  theme(
    axis.title.x=element_blank(),
    panel.grid=element_blank(),
    legend.position='none',
    axis.text.x=element_text(size=12),
    axis.title.y=element_text(size=14)
  )
```
